TO FIND

### 2013-05-27 ms.tex
Discussion

Our objective in this paper was to review the main assumptions of three
metacommunity models and illustrate how the implementation of more realistic
landsapes could reveal the importance of dispersal on community structure. We
argued in the introduction that neutral theory is useful both to understand and
predict the impact of dispersal on community organization. The review of the
different models shows that the fundamental difference between a neutral model
and the patch dynamics model is the effect of competitive interactions on
distribution, while the difference between the neutral and the species sorting
models is the effect of unequal competitive interactions. The neutral model is
thus a usefull tool to understand the joint effects of dispersal and community
interactions. Our comparison of the distribution of alpha diversity was
particularly meaningful in that respect. The simulation results show that
competitive interactions buffer the potential impact of landscape structure. The
strongest centrality-species richness relationship was observerd for the patch
dynamics, a model without any interactions. On the other hand, strong and
unequal competitive interactions minimized the effect of centrality. Our model
analysis greatly illustrates the growing recognition in metacommunity ecology
that we must move toward more realistic landscapes \parencite{Gilarranz2012}.
For field ecologists, and particularly limnologists, our review emphasize that
we need to go beyond geographic based analysis of beta-diversity (e.g.
\parencite{Legendre2005}) to topological based analyses \parencite{Peterson2013}.

@@

We introduced this article arguing that neutral theory could be used as an
instrument to predict species distribution in spatially heterogeneous
landscapes. So far we have treated only theoretical models, but we could also
envision to parameterize them and simulate real landscapes. The recruitment
probabilities defined above could be used as satistical models (likelihood
functions) to fit to empirical data. Prior information could be used to define
apriori dispersal kernels and then fit the model as in \textcite{Gravel2008}.
The fit of metapopulation models to spatially explicit landscapes was pionneered
by \textcite{Hanski1998} and recently extended to species distribution models
including both species sorting and dispersal limitations
\parencite{Boulangeat2012}. Given the parameterization, one could run neutral
models to generate null hypotheses that could be eventually compared to observed
distribution. This would make a significant improvement over traditional null
models in ecology (Gotelli1996) where there are no interactions and no dispersal
limitations.

The multivariate variance partitioning framework originally proposed by Borcard
et al. (1992) and further developped by Borcard and Legendre (2002) has been
widely used to quantify the relative importance of species sorting and dispersal
limitations in species distribution. This framework was originally proposed to
model species distribution as a function of environmental variables, taking into
accound the spatial autocorrelation of species distriution (Leduc1992,
Borcard1992, Legendre1993). This methodology has been widely used over the last
decade as a test of the neutral theory, its underlying assumption and a
quantification of dispersal limitations (e.g. Svenning2004; Hardy2004;
Gilbert2004; Cottenie2005). One problem of this approach is however that it
makes a weak test of neutrality (McGill2003), based on the description of
spatial community structure, rather than hypothesis testing. The different
models we reviewed in this article could be better employed if used to generate
expectations based on different hypotheses and then compare them. But most of
all, parameterized spatially explicit neutral models could be more useful if
used to generate expectations. For instance, neutral models could be used to
predict the consequences of habitat destruction, fragmentation or a change in
the connectinity matrix \parencite{Hubbell2008}. The spatially explicit
description of the landscape is a major improvement toward that end, providing
much flexibility in the scenarios that could be explored.

%PDP
Working with more realistic landscapes

%TP
Finally, our analyses emphasize the need to expand on the canonial neutral
theory. As pointed out by Wootoon2005Nature433, most of the unexplained
deviation of empirical community from the prediction of accurately calibrated
neutral models can be attirbuted to non-competitive interactions. Canard2012
proposed that neutral processes can explain the network structure of trophic
interactions with a good accuracy. Incorporating reasonable complexity in the
mechanisms adressed by neutral models is not a theoreticians' exercice: it will
re-inforce the usefulness of the neutral theory as an operational concept,
specifically one that can be used to derive baseline predictions about (i) the
expected local species richness, and (ii) the expected species pool
dissimilarity at the between-site and regional scales. These predictions are the
benchmark against which empirical relev\'es of species richness and community
structure ought to be compared, and coming up with realistic parameters to
calibrate these models calls for a closer cooperation and dialogue between
theoreticians and empiricists.

>>>>

Our objective in this paper was to review the main assumptions of three
metacommunity models and illustrate how the implementation of more realistic
landscapes could reveal the importance of dispersal on community structure. We
argued in the introduction that neutral theory is useful both to understand and
predict the impact of dispersal on community organization. The review of the
different models shows that the fundamental difference between a neutral model
and the patch dynamics model is the effect of competitive interactions on
distribution, while the difference between the neutral and the species sorting
models is the effect of unequal competitive interactions. The neutral model is
thus a usefull tool to understand the joint effects of dispersal and community
interactions. Our comparison of the distribution of alpha diversity was
particularly meaningful in that respect. The simulation results show that
competitive interactions buffer the potential impact of landscape structure. The
strongest centrality-species richness relationship was observed for the patch
dynamics, a model without any interactions. On the other hand, strong and
unequal competitive interactions minimized the effect of centrality. Our model
analysis greatly illustrates the growing recognition in metacommunity ecology
that we must move toward more realistic landscapes \parencite{Gilarranz2012}.
For field ecologists, and particularly limnologists, our review emphasize that
we need to go beyond geographic based analysis of beta-diversity (e.g.
\parencite{Legendre2005}) to topological based analyses \parencite{Peterson2013}.

@@

We introduced this article arguing that neutral theory could be used as an
instrument to predict species distribution in spatially heterogeneous
landscapes. So far we have treated only theoretical models, but we could also
envision to parametrize them and simulate real landscapes. The recruitment
probabilities defined above could be used as statistical models (likelihood
functions) to fit to empirical data. Prior information could be used to define
apriori dispersal kernels and then fit the model as in \textcite{Gravel2008}.
The fit of metapopulation models to spatially explicit landscapes was pioneered
by \textcite{Hanski1998} and recently extended to species distribution models
including both species sorting and dispersal limitations
\parencite{Boulangeat2012}. Given the parametrization, one could run neutral
models to generate null hypotheses that could be eventually compared to observed
distribution. This would make a significant improvement over traditional null
models in ecology (Gotelli1996) where there are no interactions and no dispersal
limitations.

The multivariate variance partitioning framework originally proposed by Borcard
et al. (1992) and further developed by Borcard and Legendre (2002) has been
widely used to quantify the relative importance of species sorting and dispersal
limitations in species distribution. This framework was originally proposed to
model species distribution as a function of environmental variables, taking into
account the spatial autocorrelation of species distribution (Leduc1992,
Borcard1992, Legendre1993). This methodology has been widely used over the last
decade as a test of the neutral theory, its underlying assumption and a
quantification of dispersal limitations (e.g. Svenning2004; Hardy2004;
Gilbert2004; Cottenie2005). One problem of this approach is however that it
makes a weak test of neutrality (McGill2003), based on the description of
spatial community structure, rather than hypothesis testing. The different
models we reviewed in this article could be better employed if used to generate
expectations based on different hypotheses and then compare them. But most of
all, parametrized spatially explicit neutral models could be more useful if
used to generate expectations. For instance, neutral models could be used to
predict the consequences of habitat destruction, fragmentation or a change in
the connectivity matrix \parencite{Hubbell2008}. The spatially explicit
description of the landscape is a major improvement toward that end, providing
much flexibility in the scenarios that could be explored.

%PDP
Working with more complex representations on landscapes has several
advantages. Real landscapes are not flat geometric objects, they are highly
structured and diverse. This diverse structure has long been recognized as
a key component of diversity. In the 19th century, Moritz Wagner noted that patterns of rivers \cite{coy04} could explain how beetles diversified. Yet, to this day, the relationship between spatial structures and biodiversity is not resolved, but networks provide powerful tools to analyze landscapes and generate
testable predictions on the relationship between community assembly
and spatial structures. For example, the neutral theory predict less diversity
in isolated communities and constant speciation (regardless of isolation) \cite{eco08}. Adding the effect of gene flow changes the predictions, but
the neutral theory do not predict adaptive radiation and might thus face
problems for predicting many long-term ecological processes.

%TP
Finally, our analyses emphasize the need to expand on the canonial neutral
theory. As pointed out by Wootoon2005Nature433, most of the unexplained
deviation of empirical community from the prediction of accurately calibrated
neutral models can be attributed to non-competitive interactions. Canard2012
proposed that neutral processes can explain the network structure of trophic
interactions with a good accuracy. Incorporating reasonable complexity in the
mechanisms addressed by neutral models is not a theoreticians' exercise: it will
re-enforce the usefulness of the neutral theory as an operational concept,
specifically one that can be used to derive baseline predictions about (i) the
expected local species richness, and (ii) the expected species pool
dissimilarity at the between-site and regional scales. These predictions are the
benchmark against which empirical relev\'es of species richness and community
structure ought to be compared, and coming up with realistic parameters to
calibrate these models calls for a closer cooperation and dialogue between
theoreticians and empiricists.
